---
title: "Andrew Bowen DATA607 Project 1"
author: "Andrew Bowen"
date: "2022-09-21"
output: html_document
---
# Project 1
```{r, message=FALSE}
library('dplyr')
library('glue')
library('plyr')
library('stringr')
```

Reading in our table using `read.delim` as it is pipe-delimited. We'll skip the comment lines (i.e., "---...") and we'll also need to manipulate this dataframe so each player has one associated row with all their information.
```{r}
github_url <- "https://raw.githubusercontent.com/andrewbowen19/cunyDATA607/main/projects/project1/tournamentinfo.txt"

df <- read.delim(github_url, sep="|", comment.char="-")
df
```

```{r}
dat <- transform(df, state = c(Pair[-1], NA))
dat
```
Shifting our rows so we have all of a player's data on one line/row
```{r}
df <- df %>% 
  mutate(state = c(Pair[-1], NA)) %>%
  mutate(rating = c(Player.Name[-1], NA))
head(df, 10)
```

Only want the odd rows now that they're consolidated. Should be able to drop exces rows
```{r}
odd_rows <- seq(nrow(df)) %% 2

df <- df[odd_rows == 0,]
head(df, 10)
```
Extracting the player's rating from the `rating` column.
```{r}
df<- df %>%
  mutate(rating = str_extract(df$rating, "R:\\s(....)"))
# Getting rid of 'R: ' prefix
df <- df %>%
  mutate(rating = gsub("R: ", "", df$rating))
```
```{r}
# Dropping unnecessary 'X' column
df <- select(df, -X, -Round)
```


Extracting the player opponent numbers -- need to remove the numbers are we don't necessarily care about wins, losses, etc. Also coercing the `rating` column to an Int type so we can average the opponent's ratings for each player.
```{r}

for (i in 1:6){
  df <- df %>%
          mutate(across(glue('Round.{i}'), str_replace,"(W|L|B|D|H|U)\\s+", '')) %>%  # Extract oppponent index
          mutate(rating = as.integer(rating)) # Coerce ratings to int
  
}

head(df, 10)
```

Looking up opponents' ratings for each player. Could likely vectorize this functionality to apply on a column basis.
```{r}
# Iterate over each player, need to grab each opponent's rankings
for (i in 1:nrow(df)) {
  # Need to calculate these metrics for each player
  rating_sum <- 0
  n_games_by_player <- 0 
  avg_rating <- 0
  
  # Then need to average against # of games player (which won't be 6 for each player)
  for (j in 1:6) {
    if (is.na(df[i, j])) {
      print("No game played.")
    }
    else {
      print(opponent_num)
      n_games_by_player <- n_games_by_player + 1
      rating_sum <- rating_sum + df[opponent_num, 'rating']

    }
  }
  # Calculating average rating of opponents based on # of games played
  avg_rating <- rating_sum / n_games_by_player
  print(glue("Avg rating: {avg_rating}"))
  df$opp_rating[i] <- avg_rating

}


head(df, 10)
```

Only selecting columns we weant and renaming them.
```{r}
df <- df %>% 
    select(Pair, Player.Name, state, Total, rating, opp_rating) %>%
    plyr::rename(c("Pair"="no",
                   "Player.Name" = "name",
                   "Total" = "total",
                   "rating" = "pre_rating",
                   "opp_rating" = "avg_opp_rating"))

head(df)
```

```{r}
write.csv(df, "elo-ratings-cleaned.csv")
```