---
title: "Andrew Bowen Data 607 Project 2"
author: "Andrew Bowen"
date: "2022-10-02"
output: html_document
---

- *Dataset 1*: [Gross Domestic Product (GDP)](https://apps.bea.gov/iTable/iTable.cfm?reqid=19&step=3&isuri=1&nipa_table_list=5&categories=survey)
- *Dataset 2:* [NFL Standings](https://www.pro-football-reference.com/years/2022/#team_stats)
- *Dataset 3:* [Gun Violence Incidents](https://www.kaggle.com/datasets/jameslko/gun-violence-data?resource=download)

---
# Dataset 1: GDP Table

Interested to see the breakdown of GDP and its different characteristics in each category. I'd like to clean this table so each observation (year & quarter) form an observation (row) within our dataset.
```{r setup, message=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
```


```{r}
gdp_data_url <- "https://raw.githubusercontent.com/andrewbowen19/cunyDATA607/main/data/gdp-wide-data.csv"
gdp <- read.csv(gdp_data_url, header=TRUE, skip=5)

head(gdp)
```

Re-naming some columns for readability
```{r}
gdp <- gdp %>%
          dplyr::rename("category" = "X",
                        "2020.Q1" = "Q1",
                        "2020.Q2" = "Q2",
                        "2020.Q3" = "Q3",
                        "2020.Q4" = "Q4",
                        "2021.Q1" = "Q1.1",
                        "2021.Q2" = "Q2.1",
                        "2021.Q3" = "Q3.1",
                        "2021.Q4" = "Q4.1",
                        "2022.Q1" = "Q1.2",
                        "2022.Q2" = "Q2.2",
                        
                        )

head(gdp)
```

```{r}
gdp$category
```

Need to transpose the dataframe in order to get it into "long format"
```{r}
gdp <- gdp %>% 
          gather(quarter, GDP, 3:12) 

gdp$category <- trimws(gdp$category, which = c("both")) # trimming whitespace
    

head(gdp)
```

Now let's make some plots. First let's get mean GDP by category
```{r}
 
gdp_avg <- gdp %>%
              group_by(category) %>%
              summarise(mean_gdp = mean(GDP))
head(gdp_avg)

```
Let's see GDP by category
```{r}
ggplot(gdp_avg, aes(x =category, y=mean_gdp)) + geom_bar(stat='identity') + theme(axis.text.x = element_text(angle = 90))
```

Let's try our hand at a "time series" and compare imorts to exports
```{r}
imp_exp <-gdp %>%
            dplyr::filter( category == 'Imports' | category == 'Exports')

ggplot(imp_exp, aes(x =quarter, y=GDP, fill=category)) + geom_point(aes(colour=factor(category)))
```


---

# Dataset 2: NFL Standings
Going to use this year's [standings tables](https://www.pro-football-reference.com/years/2022/#AFC) from pro-football-reference.com. Want to combine these into one table with all teams. We'll scrape the data directly from the page itself using the `rvest` library. These data sets also live in csv format on GitHub ([AFC](https://raw.githubusercontent.com/andrewbowen19/cunyDATA607/main/data/afc-standings.csv), [NFC](https://raw.githubusercontent.com/andrewbowen19/cunyDATA607/main/data/nfc-standings.csv)).
```{r, message=FALSE}
library(rvest)
```


```{r}
off_url <- "https://www.pro-football-reference.com/years/2022/#team_stats"
```

```{r}
# The data is read into a list of dataframes, which we'll have to unpack by conference
content <- read_html(off_url)

tables <- content %>% html_table()
tables
```

Let's combine these conference-specific dataframes into a single dataframe. While we're at it, let's filter our the sub-header rows.
```{r}
# Let's unpack the dataframes by conference
afc <- tables[[1]]
nfc <- tables[[2]]

# Combining and cleaning our two tables
df <- rbind(afc, nfc) 

df <- df %>%
        dplyr::filter(Tm != "NFC East" ,
               Tm != "AFC East" ,
               Tm != "NFC South" ,
               Tm != "AFC South" ,
               Tm != "NFC North" ,
               Tm != "AFC North" ,
               Tm != "NFC West" ,
               Tm != "AFC West"
              ) %>%
        dplyr::rename("wl_pct" = "W-L%")  # Makes it easier to access this column when renamed


df #head(df)
```
Lastly, we're not dealing with multiple `NA` values in our dataframe, so let's coerce relevant columns to numeric types for calculation later.
```{r}
df[, 2:13] <- sapply(df[, 2:13], as.factor)
```


First, I'd like to see a scatter plot of a team's point-differential ($`PF` - `PA`$) versus a team's win-loss percentage. Since no teams have had a bye week yet, using win-loss percentage is a better measure of a team's performance (in terms of winning football games).
```{r}
ggplot(df, aes(x=PD, y=wl_pct)) + geom_point()
```
It's still early in the season (4 games played at the time of this writing), so there aren't many possible records, leading to a "stepwise" pattern in the above plot.


I'd also like to see a plot of offensive and defensive SRS, which measures a team relative to average
```{r}
ggplot(df, aes(x=DSRS, y=OSRS)) + geom_point()
```
Let's see if there's any correlation between a team's defensive rating versus offensive.
```{r}
cor(as.numeric(df$OSRS), as.numeric(df$DSRS))
```


---

# Dataset 3: Gun Violence
Using this [gun violence dataset](https://www.kaggle.com/datasets/jameslko/gun-violence-data?resource=download) from Kaggle. Since the original file is too big to [store on GitHub](https://raw.githubusercontent.com/andrewbowen19/cunyDATA607/main/data/gun-violence-sample.csv), I took the first 10000 rows. Will need to tidy up some columns before we can analyze. 
```{r}
library(sf)
library(mapview)
```


```{r}
gv_url <- "https://raw.githubusercontent.com/andrewbowen19/cunyDATA607/main/data/gun-violence-sample.csv"

gv <- read.csv(gv_url)

head(gv, 15)
```

Let's start by cleaning up our `gun_stolen` and `gun_type` columns. We don't need the `0::` prefixes, for instance.

```{r}
gv <- gv %>%
        mutate(gun_type = str_replace_all(gun_type, "^[^\\|]*\\|{2}", "")) %>%
        mutate(gun_stolen = str_replace_all(gun_stolen, "^[^\\|]*\\|", ""))


head(gv, 10)
```


Let's also break up our `participant_age` column so it's a bit cleaner for later analysis
```{r}
gv <- gv  %>%
        mutate(perp_age = str_replace_all(str_extract(participant_age, "[0-9]::(\\d{1,2})"), "\\d::", ""))

head(gv)
```


Now let's make some plots to make sense of our dataset
```{r}
ggplot(gv, aes(x=n_killed)) + geom_histogram(binwidth=1)
```

Also would like to see a histogram of ages involved in gun violence incidents.
```{r}
# Need to remove NA values for a cleaner graph
gv_with_age <- gv %>%
                filter(!is.na(perp_age)) %>%
                mutate(perp_age = as.integer(perp_age))

ggplot(gv_with_age, aes(x=perp_age)) + geom_histogram(binwidth=5)
```


I'd like to see the locations (via `latitude` and `longitude`) mapped out. Using the `mapview` 
```{r}
# Need to remove NA
gv_with_locs <- gv %>%
                        filter(!is.na(longitude), !is.na(latitude))

mapview(gv_with_locs, xcol="longitude", ycol="latitude", crs = 4326)


```


