---
title: "Andrew Bowen Data 607 Project 2"
author: "Andrew Bowen"
date: "2022-10-02"
output: html_document
---

- Dataset 1: [Gross Domestic Product (GDP)](https://apps.bea.gov/iTable/iTable.cfm?reqid=19&step=3&isuri=1&nipa_table_list=5&categories=survey)
- 

```{r setup, message=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
```


```{r}
gdp_data_url <- "https://raw.githubusercontent.com/andrewbowen19/cunyDATA607/main/data/gdp-wide-data.csv"
gdp <- read.csv(gdp_data_url, header=TRUE, skip=5)

head(gdp)
```

Re-naming some columns for readability
```{r}
gdp <- gdp %>%
          dplyr::rename("category" = "X",
                        "2020.Q1" = "Q1",
                        "2020.Q2" = "Q2",
                        "2020.Q3" = "Q3",
                        "2020.Q4" = "Q4",
                        "2021.Q1" = "Q1.1",
                        "2021.Q2" = "Q2.1",
                        "2021.Q3" = "Q3.1",
                        "2021.Q4" = "Q4.1",
                        "2022.Q1" = "Q1.2",
                        "2022.Q2" = "Q2.2",
                        
                        )

gdp
```

```{r}
gdp$category
```

Need to transpose the dataframe in order to get it into "long format"
```{r}
gdp <- gdp %>% 
          gather(quarter, GDP, 3:12) 

gdp$category <- trimws(gdp$category, which = c("both")) # trimming whitespace
    

gdp
```

Now let's make some plots. First let's get mean GDP by category
```{r}
 
gdp_avg <- gdp %>%
              group_by(category) %>%
              summarise(mean_gdp = mean(GDP))
gdp_avg

```
Let's see GDP by category
```{r}
ggplot(gdp_avg, aes(x =category, y=mean_gdp)) + geom_bar(stat='identity') + theme(axis.text.x = element_text(angle = 90))
```

Let's try our hand at a "time series" and compare imorts to exports
```{r}
imp_exp <-gdp %>%
            dplyr::filter( category == 'Imports' | category == 'Exports')

ggplot(imp_exp, aes(x =quarter, y=GDP, fill=category)) + geom_point(aes(colour=factor(category)))
```


---

## Dataset 2
Going to use this year's [standings tables](https://www.pro-football-reference.com/years/2022/#AFC) from pro-football-reference.com. Want to combine these into one table with all teams. I grabbed the csv files and uploaded them to GitHub, for reproducability.

```{r, message=FALSE}
# library(XML)
# library(RCurl)
# library(htmltools)
library(textreadr)
```


```{r}
off_url <- "https://widgets.sports-reference.com/wg.fcgi?css=1&site=pfr&url=%2Fyears%2F2022%2F&div=div_team_stats"#"https://www.pro-football-reference.com/years/2022/#team_stats"
```

```{r}
df <- read_html(off_url)
df
```


---

## Dataset 3
Using this [gun violence dataset](https://www.kaggle.com/datasets/jameslko/gun-violence-data?resource=download) from Kaggle. Will need to tidy up some columns
