---
title: "DATA 607 Project 4"
author: "Andrew Bowen"
date: "2022-11-15"
output: html_document
---

```{r setup, message=FALSE}
library(stringr)
library(tensorflow)
library(tfdatasets)
library(keras)
library(glue)

# install_tensorflow()
```


I'd like to try some text classification using the `tensorflow` package in R. I'm leaning on this [tutorial page](https://tensorflow.rstudio.com/tutorials/keras/text_classification) to get some of the basics, as I've only used tensorflow in python before, not R

Loading in our [ham dataset](https://spamassassin.apache.org/old/publiccorpus/20021010_easy_ham.tar.bz2)
```{r}
# Loading ham dataset
ham_url <- "https://spamassassin.apache.org/old/publiccorpus/20021010_easy_ham.tar.bz2"

ham_dataset <- get_file(
  "data/20021010_easy_ham",
  ham_url,
  untar = TRUE,
  cache_dir = '.',
  cache_subdir = ''
)

# Loading spam dataset
spam_url <- "https://spamassassin.apache.org/old/publiccorpus/20030228_spam_2.tar.bz2"

spam_dataset <- get_file(
  "data/20030228_spam_2",
  spam_url,
  untar = TRUE,
  cache_dir = '.',
  cache_subdir = ''
)
```


```{r}
ham_dir <- file.path("easy_ham")
spam_dir <- file.path("spam_2")
```

```{r}
# Renaming unzipped files in our spam and ham dirs so they have a ,txt extension
dirs <- c("spam_2", "easy_ham")

for (d in dirs) {
  dir_path <- glue("data/{d}/")
  for (f in list.files(dir_path)) {
    if (!grepl(".txt", f)) {
      new_name <- glue("{dir_path}{f}.txt")
      file.rename(glue("{dir_path}{f}"), new_name)
    }
  }
}

```

Generating training datasets from our directories containing the raw text files. [`keras` looks for the sub-folders](https://datascience.stackexchange.com/questions/51671/keras-flow-from-directory-returns-0-images) within our data folder as the sub-classes
```{r}
seed <- 1234


raw_train_ds <- text_dataset_from_directory(
  'data/',
  batch_size = 32,
  validation_split = 0.1,
  subset = 'training',
  seed = seed
)

```


```{r}
raw_val_ds <- text_dataset_from_directory(
  'data/',
  batch_size = 32,
  validation_split = 0.1,
  subset = 'validation',
  seed = seed
)
```


```{r}
# Content-Transfer-Encoding: 7bit


```


