---
title: "Andrew Bowen Data 607 HW 6"
author: "Andrew Bowen"
date: "2022-10-24"
output: html_document
---

# Andrew Bowen Data 607 HW 6

Using data from the [NY Times public api](https://developer.nytimes.com/apis) to read JSON data into an R dataframe.

```{r setup, message=FALSE}
library(httr)
library(dplyr)
library(glue)
library(jsonlite)
library(ggplot2)
library(tidyr)
```

First we'll need to defin our API endpoint (URL). I chose the [NT Times Books API](https://developer.nytimes.com/docs/books-product/1/overview), which provides data on NY Times bestseller lists. I am setting my API key value in the environment variable `NY_TIMES_API_KEY`. Please check my blackboard submission for my key value, or feel free to use your own key as it's a public-facing API. 
```{r}

api_key = Sys.getenv("NY_TIMES_API_KEY")
url <- glue('https://api.nytimes.com/svc/books/v3/lists/current/hardcover-fiction.json?api-key={api_key}')

# Call the API and get an HTTP response
response <- GET(url, format="json")

response
```

Now we need to extract the content from that API response. We can use the`content` attribute from the `httr::response` object. We can then use the `jsonlite::fromJSON` method to parse the response into a dataframe.
```{r}
# Getting content of HTTP response
details <- content(response, as="text")

df <- jsonlite::fromJSON(details)

# extracting our books dataframe from the results field
df <- df$results$books

head(df)
```

Let's do some quick visualization of our dataset
```{r}
ggplot(df, aes(x = weeks_on_list)) + geom_histogram()
```



