---
title: "Andrew Bowen DATA607 HW4"
author: "Andrew Bowen"
date: "2022-09-27"
output: html_document
---
## Andrew Bowen - DATA607 HW 4


```{r setup, message=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
```

## Question 1
I manually typed the dataset into a .csv file via vim which I then pushed to GitHub. I kept the formatting of the table to reflect the 'wide' structure it was presented in. Reading into our R dataframe from the remote repo.
```{r}
data_url <- "https://raw.githubusercontent.com/andrewbowen19/cunyDATA607/main/data/flight-arrival-delays.csv"
df <- read.delim(data_url, header=TRUE, sep=",")

df
```


Removing/re-formatting some of the missing values in our raw dataset.
```{r}
# Removing empty row with NA values
df <- df %>%
      drop_na(Los.Angeles, Phoenix, San.Diego, San.Fransisco, Seattle)
      
# Filling in missing carrier vals
df[2, "Carrier"] <- "Alaska"
df[4, "Carrier"] <- "AMWEST"
df
```

## Question 2

It looks like we'd be able to describe this dataset across 4 key dimensions: 
  - Airline (ALASKA or AMWEST)
  - Status (on time vs delayed)
  - Destination (e.g, Los Angeles, Phoenix, etc.)
  - Arrival Delay Count (how many flights arrived at those airports)
  
Using [tidyr's `gather`](https://www.statology.org/gather-function-in-r/) function which transforms our dataset into key-value pairs across multiple columns. This transforms our dataframe into a nice, clean 'long' format which we can use more easily in our analysis
```{r}
df <- tidyr::gather(df, "dest", "arrival_delays", 3:7)

df
```
```{r}
# Renaming our columns to snake_case
df <- df %>% rename(carrier=Carrier, status=Status, destination = dest)
```

```{r}
library(stringr)
# Replacing city names with proper spelling
df <- df %>%
  mutate(across('destination', str_replace, "\\.", " "))

df
```



```{r}
# Let's see the total delays for each airline
delays_by_airline <- df %>%
  filter(status == "delayed") %>%
  group_by(carrier) %>%
  summarise(
            total_delays=sum(arrival_delays)
          )

delays_by_airline
```

Plotting the delayed arrivals by destination
```{r}
delays <- df %>% filter(status == "delayed") %>%
              group_by(destination) %>%
              summarise(total_delays = sum(arrival_delays))


ggplot(delays, aes(x=destination, y=total_delays)) + geom_bar(stat='identity')# + geom_text(aes())
```
Plotting the all arrivals (delayed and on time) for all flights across each airline.
```{r}
delays_by_airline_status <- df %>%
                        group_by(carrier, status) %>%
                        summarise(total_delays = sum(arrival_delays))
delays_by_airline_status
ggplot(delays_by_airline_status, aes(x=interaction(carrier, status), y=total_delays, fill=factor(carrier))) + geom_bar(stat='identity')
```

