---
title: "Andrew Bowen Data 607 HW 5"
author: "Andrew Bowen"
date: "2022-10-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## TODO:
- Ensure "file paths" are github urls
- HTML and XML read-in
  
Using a few libraries to parse the various data formats
```{r message=FALSE}
library(XML)
library(xml2)
library(rjson)
library(dplyr)
library(tidyr)
```

First, let's lay out where we're
```{r}
# TODD: replace local paths with GitHub URL on better network conn
json_url <- "~/CUNY/cunyDATA607/data/books.json"
html_url <- "~/CUNY/cunyDATA607/data/books.html"
xml_url <- "~/CUNY/cunyDATA607/data/books.xml"
```

## JSON read-in
```{r}
json_data <- fromJSON(file=json_url)  # Replace this with github url

json_df <- as.data.frame(json_data)

json_df
```
This table looks pretty neatly formatted. Since we want one "item"/"observation" to constitute a row in our books table, it makes sense that 

---
## XML Read-in

We put the same data into our XML file, but in a different format. We still want each book to constitute a row in our dataframe
```{r}
xml_data <- xml2::read_xml(xml_url)

# Parse XML and convert into R dataframe
book_xml <- XML::xmlParse(xml_data)
xml_df <- XML::xmlToDataFrame(book_xml)

xml_df
```

Let's convert our rating column to a double, to be consistent with our `json_df`
```{r}
xml_df$rating <- as.double(xml_df$rating)
```

---
## HTML read-in

Going to use the `XML` library from aboce and its `readHTMLTable` method in order to grab data from out html table.
```{r}
# Read in using XML::readHTMLTable method
html <- readHTMLTable(html_url, as.data.frame=TRUE)
html_df <- as.data.frame(html)

html_df
```

This dataframe is close to the ones we generated vai XML and JSON, with the `rating` type as a `character` rather than an `double`, and the column names prefixed by *NULL.*. Let's 

```{r}
html_df <- html_df %>% 
                  rename("title" = "NULL.title",
                         "author" = "NULL.author",
                         "ISBN" = "NULL.ISBN",
                         "rating" = "NULL.rating",
                         "genre" = "NULL.genre")
html_df$rating <- as.double(html_df$rating)

html_df
```


* * *