---
title: "Andrew Bowen DATA 607 HW 3"
author: "Andrew Bowen"
date: "2022-09-13"
output: html_document
---
# Homework Assignment 3: DATA607

```{r, results='hide'}
library('dplyr')
library('tidyverse')
```

## Question 1
Using [FiveThirtyEight's College major list](https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv) csv file, pulling the data into our `majors` dataframe straight from GitHub for reproducibility.
```{r}
major_list_url <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv"
majors <- read.csv(major_list_url, header=TRUE, sep=",")

head(majors, 10)
```
Using the `filter` function that comes with the `dplyr` package
```{r}
dplyr::filter(majors, grepl("DATA|STATISTICS", Major))
```
## Question 2
```{r}
initial_str <- '[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"
[9] "elderberry"   "lime"         "lychee"       "mulberry"
[13] "olive"        "salal berry"'


st <- str_extract_all(initial_str, "(bell pepper|bilberry|blackberry|blood orange|blueberry|cantaloupe|chili pepper|cloudberry|elderberry|lime|lychee|mulberry|olive|salal berry)")
#st <- unlist(str_remove_all(initial_str, "\\s|\\[.\\]"))

extracts <- unlist(st)
extracts
```

## Question 3
Describe, in words, what these expressions will match:

`(.)\1\1`
This expression will match for strings that have the same characterer repeated two times in a row (for example: l`oo`k). This comes from the double backreference `\1` refering to the capture group `(.)`
```{r}
pattern <- "(.)\1\1"
# str_view(fruit, pattern, match=TRUE)
```

`"(.)(.)\\2\\1"`
This regex will match to any character groups that feature a 2-character group followed by it's "mirror" (e.g., `arra`y).
```{r}
# str_view(fruit, "(.)(.)\\2\\1")
```

`(..)\1`
This regex will match to a word with a 2-character sequence that is repeated
```{r}
# str_view("FOFO", "(..)\1")
```


`"(.).\\1.\\1"`
This regex will match to any letter that appears 3 times (hence the 2 back-references) within a word.

```{r}
# str_view(fruit, "(.).\\1.\\1")
```
`"(.)(.)(.).*\\3\\2\\1"`
This regex matches to strings that have a 3-character sequence that is "mirrored" later in the word. For example `rac`e`car`

```{r}
# str_view(fruit, "(.)(.)(.).*\\3\\2\\1")
```

---
## Question 4
Match the first and last letter of a word if they are the same:
```{r}
test_str <- c("abba", "dancing", "queen", "waterloo", "chiquitita")

str_view(test_str, "^(.).*\\1$")# "^(.).?\1$")
```


Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
```{r}
# The letters don't have to be the same (i.e. 'ch'), so we can use a character range
str_view(test_str, "([a-zA-Z][a-zA-z]).*\\1")
```

Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)
```{r}
# Using two capture group references to catch the same character two additional times
str_view(test_str, "(.).*\\1.*\\1")
```





