---
title: "Andrew Bowen Data 607 HW 5"
author: "Andrew Bowen"
date: "2022-10-13"
output: html_document
---

*TODO:*
- [x] Ensure file paths are github urls
- [x] HTML and XML read-in
- Publish to Rpubs
  
Using a few libraries to parse the various data formats
```{r message=FALSE}
library(rvest)
library(XML)
library(xml2)
library(rjson)
library(dplyr)
library(tidyr)
```

First, let's lay out where we're
```{r url-def}
json_url <- "https://raw.githubusercontent.com/andrewbowen19/cunyDATA607/main/data/books.json"
html_url <- "https://raw.githubusercontent.com/andrewbowen19/cunyDATA607/main/data/books.html"
xml_url <- "https://raw.githubusercontent.com/andrewbowen19/cunyDATA607/main/data/books.xml"
```

## JSON read-in
```{r json-read-in}
json_data <- fromJSON(file=json_url, simplify=TRUE)

json_df <- as.data.frame(json_data)

json_df
```
This table looks pretty neatly formatted. Since we want one "item"/"observation" to constitute a row in our books table, it makes sense that 


## XML Read-in

We put the same data into our XML file, but in a different format. We still want each book to constitute a row in our dataframe
```{r xml-read-in}
xml_data <- xml2::read_xml(xml_url, as_html=FALSE)

# Parse XML and convert into R dataframe
book_xml <- XML::xmlParse(xml_data)
xml_df <- XML::xmlToDataFrame(book_xml)

xml_df
```

Let's convert our rating column to a double, to be consistent with our `json_df`
```{r xml-typing}
xml_df$rating <- as.double(xml_df$rating)
```


## HTML read-in

Going to use the `XML` library from above and its `readHTMLTable` method in order to grab data from out html table.
```{r html-read-in}


html_df <- (rvest::read_html(html_url) %>% html_table)[[1]]

# Read in using XML::readHTMLTable method
# html <- readHTMLTable(html_url), as.data.frame=TRUE) # "../data/books.html
# html_df <- as.data.frame(html)

html_df
```

This dataframe is close to the ones we generated vai XML and JSON, with the `rating` type as a `character` rather than an `double`, and the column names prefixed by *NULL.*. Let's 

```{r html-standrdize}

html_df <- (rvest::read_html(html_url) %>% html_table)[[1]]

html_df$rating <- as.double(html_df$rating)

html_df
```
This looks consistent with our above XML and JSON dataframes!
