---
title: "DATA 607 Final Project"
author: "Andrew Bowen"
date: "2022-11-23"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, echo=FALSE, message=FALSE}
library(httr)
library(jsonlite)
library(tidyverse)
library(sf) # for geospatial data lookup
library(glue)
```
## Motivation
- **Research Question**: Do environmental factors (like air quality) impact a student's academic performance?

## Datasources
- [AQ Air Quality Public REST API](https://docs.openaq.org/reference/introduction-1)
- [Open Weather Map API](https://openweathermap.org/api)
- [NYC Open Data Education files on test results](https://data.cityofnewyork.us/browse?category=Education&q=Test&sortBy=relevance) : while not a perfect indicator of academic performance, these do give some insight into the academic progress of students in different areas of the city.

## Academic Performance Data
- [Dataset from NYC Open Data](https://data.cityofnewyork.us/Education/2006-2012-Math-Test-Results-All-Students/3mrr-8h5c/data) (csv) -- test results by student.
- Can use District Borough Number (DBN) to lookup geographic locations and compare against AQ data
```{r, echo=TRUE}
dbn_url <- "https://raw.githubusercontent.com/andrewbowen19/cunyDATA607/main/projects/final_project/2006_-_2012__Math_Test_Results__-_All_Students.csv"

tests <- read.csv(dbn_url)

head(tests)
```

## Getting School Geospatial Data
Looking up [school district GIS data from NYC Open Data](https://data.cityofnewyork.us/Education/School-Districts/r8nu-ymqj). School point locations can be found [here](https://data.cityofnewyork.us/Education/School-Point-Locations/jfju-ynrr), we'll download them and unzip the file to get the shp file
```{r}
# Create temporary files
temp <- tempfile()
temp2 <- tempfile()

# Download the zip file and save to 'temp' 
school_locs_url <- "https://data.cityofnewyork.us/download/jfju-ynrr/application%2Fzip"
download.file(school_locs_url, temp)

# Unzip the folder contents and then read in the shp file to a dataframe
unzip(zipfile = temp, exdir = temp2)
school_locs <- sf::read_sf(temp2)
```
## Cleaning School Location Data
```{r}
# Cleaning and re-naming the ATS_CODE column to 
school_locs <- school_locs %>% 
                  dplyr::rename( "DBN"="ATS_CODE")
school_locs$DBN <- str_replace_all(school_locs$DBN, "\\s+", "")

# Projected CRS: NAD83
school_locs_sf <- st_as_sf(school_locs, coords=c("YCoord", "XCoord"), NAD83)

school_locs <- st_transform(school_locs_sf, crs = 4326 ) %>%
      mutate( lat = st_coordinates(.)[,2],lon = st_coordinates(.)[,1])
```

## Joining & Cleaning Academic Data
Let's join our `school_locs` dataframe to our `tests` 
```{r}
#  Adding in borough lookup as well
school_boro_lookup = data.frame(BORO=c("K", "M", "Q", "R", "X"),
                                borough=c("Brooklyn", "Manhattan", "Queens", "Staten Island", "Bronx")
                                )
school_data <- as.data.frame(left_join(tests, school_locs, by="DBN")) %>%
                            left_join(., school_boro_lookup, by="BORO") %>%
                            rename("year" = "Year") %>%
                            mutate(year = as.character(year))
```

## Pulling Air Quality Data for New York City

Looking up air quality for school locations in [New York's geographical area](https://www.latlong.net/place/new-york-city-ny-usa-1848.html).




## Air quality data
Using this [NYC Open Data Air Quality Dataset from their API](https://data.cityofnewyork.us/Environment/Air-Quality/c3uy-2p5r). Going to call their API


Breaking down air quality data by borough. Air pollution [varies borough-to-borough in New York City](https://vega.github.io/schema/vega-lite/v4.json) so we'll group our data that way. We'll need to add a `borough` column to our Air quality dataset
```{r}
#  Getting data from API call
# https://data.cityofnewyork.us/Environment/Air-Quality/c3uy-2p5r
nyc_open_data_aq_url <- "https://data.cityofnewyork.us/resource/c3uy-2p5r.json"
nyc_aq <- as.data.frame(fromJSON(nyc_open_data_aq_url))

# Adding in borough lookup code to compare to school data
# Also adding year column to compare with testing data
borough_lookup <- data.frame(borough_code = c("1", "2", "3", "4", "5"),
                    borough=c("Bronx",
                    "Brooklyn",
                    "Manhattan",
                    "Queens",
                    "Staten Island"))
nyc_aq <- nyc_aq %>%
            mutate(borough_code = str_extract(geo_join_id, "^\\d"),
                   year = str_extract(start_date, "^\\d{4}")) %>%
            left_join(., borough_lookup, by="borough_code")


head(nyc_aq)
```
## Additional AQ Data Cleaning
```{r}
nyc_aq <- nyc_aq %>% pivot_wider(., names_from = name, values_from=data_value) %>%
            rename("ozone" = "Ozone (O3)", "sulfur_o2" = "Sulfur Dioxide (SO2)") %>%
            mutate(ozone = as.double(ozone), sulfur_o2 = as.double(sulfur_o2))
```


## Joining our Air Quality and Academic
```{r}
df <- left_join(school_data, nyc_aq, by=c("borough", "year"))
```

```{r, echo=FALSE, message=FALSE}
# Writing to CSV file to avoid future API calls in our analysis
write.csv(df, "school_aq_data.csv")
```

## Data Analysis

Let's get a sense of the distributions for our `ozone` and `sulfer_o2` variables
```{r dist-ozone-so2}
bin_width = 0.5
# Ozone distribution
ggplot(df, aes(x=ozone)) + geom_histogram(binwidth=bin_width)

# SO2 distribution
ggplot(df, aes(x=sulfur_o2)) + geom_histogram(binwidth=bin_width)
```


## ANOVA ()
Running an Analysis of Variance (ANOVA) hypothesis test to see if there's a statistically significant different in ozone levels between boroughs.We'll conduct this test with significance level ($\alpha = 0.05$) and the following hypotheses:

- $H_0$: There is **no** difference in mean ozone (O3) levels between NYC boroughs

- $H_a$: There **is a** difference in mean ozone (O3) levels between NYC boroughs

```{r anova-ozone-by-borough}
ozone_aov <- aov(ozone ~ borough, data = df )
summary(ozone_aov)
```

## ANOVA (SO2)

- $H_0$: There is **no** difference in mean sulfur dioxide (SO2) levels between NYC boroughs

- $H_a$: There **is a** difference in mean sulfur dioxide (SO2) levels between NYC boroughs

```{r anova-so2-by-borough}
so2_aov <- aov(sulfur_o2 ~ borough, data = df )
summary(so2_aov)
```

Making a choropleth map of each borough, colored by mean `ozone` and `sulfur_o2` value, with school locations plotted.
```{r, message=FALSE, echo=FALSE}
library(maptools)
library(ggmap)
library(tigris)
library(rgdal)
library(broom)

r <- GET('http://data.beta.nyc//dataset/0ff93d2d-90ba-457c-9f7e-39e47bf2ac5f/resource/35dd04fb-81b3-479b-a074-a27a37888ce7/download/d085e2f8d0b54d4590b1e7d1f35594c1pediacitiesnycneighborhoods.geojson')
nyc_neighborhoods <- readOGR(content(r,'text'), 'OGRGeoJSON', verbose = F)

nyc_neighborhoods_df <- broom::tidy(nyc_neighborhoods)

# Plotting schools colored by ozone levels
ggplot() + 
  geom_polygon(data=nyc_neighborhoods_df, aes(x=long, y=lat, group=group)) + 
  geom_point(data=df, aes(x = lon, y=lat, colour=ozone))
```

```{r}
ggplot() + 
  geom_polygon(data=nyc_neighborhoods_df, aes(x=long, y=lat, group=group)) + 
  geom_point(data=df, aes(x = lon, y=lat, colour=sulfur_o2, fill=sulfur_o2))
```

